---
title: "project Explore and Summarize Data"
output: html_document
author: "Dalal"
date: "16th January 2019"
---
========================================================

# Data analysis and exploration

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.


library(ggplot2)
library(gridExtra)

```

```{r echo=FALSE, Load_the_Data}

# Load the Data
getwd() # know the path 
df <- read.csv("wineQualityReds.csv") # load the data 
df$X <- NULL # delete the first variable because it is just index 
names(df)

```

# Introduction
in the Explore and Summarize Data project i will explore Red wine Data set, 
the main objective of this project is to explore the chemical variables that have impact on the wine 
this data set contain 12 variables and 1599 observations.

# Univariate Plots Section


```{r echo=FALSE}

summary(df) # statistic information about the Red wine dataset

```

From the summary i see that the Red wines have:

Chlorides: between 0.012 and 0.611 and the Mean value = 0.087
PH: between 2.740 and 4.010 and the Mean value = 3.311
alcohol: between 8.40 and 14.90 and the mean value = 10.42
Quality: between 3.000 and 8.000 and the Mean value = 5.636



```{r echo=FALSE}

qplot(factor(quality), data=df, xlab="Quality") + theme_bw() +
  xlab(("Quality")) + ylab("Wine count")

```


The quality of red wine is normally distributed  around 5,
thats mean the quality of red wine collection is good 

```{r echo=FALSE}

count_c <- function(var, bins) {
  ggplot(aes_string(x = var), data = df) +
    geom_histogram( binwidth = bins)
}

p1 <- count_c("fixed.acidity", 0.5)
p2 <- count_c("volatile.acidity", 0.05)
p3 <- count_c("citric.acid", 0.03)
p4 <- count_c("residual.sugar", 0.5)
p5 <- count_c("chlorides", 0.025)
p6 <- count_c("free.sulfur.dioxide", 5)
p7 <- count_c("total.sulfur.dioxide", 10)
p8 <- count_c("density", 0.001)
p9 <- count_c("pH", 0.05)
p10 <- count_c("sulphates", 0.1)
p11 <- count_c("alcohol", 0.25)

grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, ncol = 4)

```

To see all the chemical variables that have the impact on the wine, and i fount that (residual sugar, chlorides, free sulfur dioxide , total sulfur dioxide, and sulphates) are positive skew and the (density and PH ) are normally distributed.

```{r echo=FALSE}

p1 <- count_c("free.sulfur.dioxide", 5)
p2 <- count_c("free.sulfur.dioxide", 0.5) +
  scale_x_log10()
p3 <- count_c("total.sulfur.dioxide", 10)
p4 <- count_c("total.sulfur.dioxide", 0.10) +
  scale_x_log10()
grid.arrange(p1, p2, p3, p4,ncol = 2)

```

i grouped the free and total sulfur dioxide together and, 
from the histogram above i see that both free and total sulfur.dioxide
have normal distributions.

```{r echo=FALSE}

p1 <- count_c("fixed.acidity", 0.5) 
  scale_x_log10()
p2 <- count_c("volatile.acidity", 0.05) +
  scale_x_log10()
p3 <- count_c("citric.acid", 0.03) 
  scale_x_log10()
grid.arrange(p1, p2, p3, ncol = 2)


```

i grouped the acids together and,
from the histogram above i see that:
The fixed acidity of red wine between 5 to 11 
and
The Volatile acidity of red wine between 0.2 - 0.8
and
The citric acid of red wine between 0.01 - 0.50


```{r echo=FALSE}

ggplot(df, aes(x=alcohol)) + 
  geom_density()  +
  xlab(("Alcohol %")) + ylab("density")


```

from the chart above it seems that the alcohol content follow an abnormal distribution and it contains a high peak at the lower.




```{r echo=FALSE}

#define a new variable alcohol density
df$alcohol.density <- ifelse (df$alcohol <10.5,'Low Alcohol' , ifelse(
  df$alcohol < 12.5 ,'Medium Alcohol', 'High Alcohol'))

df$alcohol.density <- ordered (df$alcohol.density ,
                                levels=c('Low Alcohol','Medium Alcohol','High Alcohol'))
    ggplot(df, aes(alcohol.density)) +
    geom_bar(aes()) +
    ylab("wine count")

```


From the histogram above i define a new variable calles alcohol density 
depend on the alcohol to see the density of the alcohol 
and see the highest density 
low Alcohol with alcohol between (8.3 to 10.5)
Medium Alcohol with alcohol between (10.55 to 12.5 )
high Alcohol with alcohol between (12.6 to 14.9)
and i found that low has the highest Alcohol the count is around 1000 and 
then Medium has Alcohol around 550 and High has Alcohol around 60.

```{r echo=FALSE}

#define a new variable alcohol quality
df$alcohol.quality <- ifelse (df$quality <5, 'low' , ifelse(
  df$quality < 7 ,'Medium', 'v.good'))

df$alcohol.quality <- ordered (df$alcohol.quality , levels=c('low','Medium','v.good'))
    ggplot(df, aes(alcohol.quality)) +
    geom_bar(aes()) +
  ylab("Wine count")
    
```

From the histogram above i define a new variable calles alcohol quality depend on quality 
low quality is < 5 
medium < 7
v.good > 7 
to see the highest average alcohol quality and found that is medium is the highest 


# Univariate Analysis

### What is the structure of your dataset?

```{r echo=FALSE}
str(df)
```

wine Dataset contain 1599 observation and 12 variables (fixed acidity - 
volatile acidity - citric acid - residual sugar chlorides - free sulfur dioxide -
total sulfur dioxide - density - pH - sulphates - alcohol- quality ) 

and i create 2 variables alcohol category and alcohol quality.

### What is/are the main feature(s) of interest in your dataset?

The wine Quality and Alcohol is the main features. 

### What other features in the dataset do you think will help support your \ investigation           into your feature(s) of interest?

sulphates and density.

### Did you create any new variables from existing variables in the dataset?

yes, alcohol density and alcohol quality

### Of the features you investigated, were there any unusual distributions? \ Did you perform any operations on the data to tidy, adjust, or change the form \ of the data? If so, why did you do this?

yes i observed some unusual distribution with the fixed acidity, citric acid,
volatileacidity, free sulfur dioxide and total sulfur dioxide variables and 
i use log10 and to understand the distribution better.



# Bivariate Plots Section


```{r echo=FALSE}

# fixed acidity with density

ggplot(df, aes(fixed.acidity, density)) + 
  geom_jitter(alpha=1/5) +  
  stat_smooth(method = "glm", formula = y ~ x) +
  xlab(("fixed.acidity ")) + ylab("density")

```

from the chart above i see that fixed acidity is increasing with density. 

```{r echo=FALSE}

# density with alcohol
ggplot(df, aes(density, alcohol)) + 
  geom_jitter(alpha=1/5) + 
  stat_smooth(method = "glm",formula = y ~ x)+
  xlab(("Alcohol % ")) + ylab("density")

```

from the chart above i see that density is increasing while the alcohol decreasing 


```{r echo=FALSE}

# fixed acidity with PH

ggplot(df, aes(fixed.acidity, pH)) + 
  geom_jitter(alpha=1/5) + 
  stat_smooth(method = "glm", formula = y ~ x)

```

from the chart above i see that the relationship is  negative,
lower pH correlates with higher fixed acidity 


```{r echo=FALSE}

ggplot(df, aes(density,  residual.sugar )) + 
  geom_jitter(alpha=1/5) + 
  stat_smooth(method = "glm", formula = y ~ x)

```

from the chart above i see that residual sugar is increasing with density. 

```{r echo=FALSE}

ggplot(aes(x = volatile.acidity, y = citric.acid), data = df) +
  geom_point(alpha=1/5) +
  geom_smooth(se = FALSE)
```

From the chart above i see that there is negative correlation between:
citric acid and volatile acidity

```{r echo=FALSE}

ggplot(aes(x = alcohol.density, y = density), data = df) +
  geom_boxplot()

```

from the chart above i see that the wine with High alcohol
has the lowest meadian density.


```{r echo=FALSE}

ggplot(aes(x = alcohol.quality, y = volatile.acidity), data = df) +
  geom_boxplot()

```

from the chart above i see that the wine with the v.good quality
has the lowest meadian volatile acidity.

```{r echo=FALSE}
ggplot(aes(x = alcohol.quality, y = alcohol), data = df) +
  geom_boxplot()
```



# Bivariate Analysis



### Talk about some of the relationships you observed in this part of the \ investigation. How did the feature(s) of interest vary with other features in \the dataset?

I see there is a Positive correlation between:
(fixed acidity with density)  and (residual sugar with density).
and Negative correlation between:
(density with alcohol) and (fixed acidity with PH)

From the box plot, i see that there is a negative correlation between:
(Alcohol quality with volatile.acidity).
and positive correlation between:
(Alcohol quality with alcohol).

### Did you observe any interesting relationships between the other features \(not the main feature(s) of interest)?

negative correlation between (citric acid and volatile acidity)
positive relationship between (fixed.acidity and density)

### What was the strongest relationship you found?

alcohol quality with alcohol.

# Multivariate Plots Section



```{r echo=FALSE}

ggplot(aes(x = chlorides, y = free.sulfur.dioxide), data = df) +
  geom_point(aes(color = alcohol.density),
             position = position_jitter(h = 0), alpha = 1/5) +
  scale_colour_brewer(palette = "RdPu")  + theme_dark() +
  scale_x_log10() +
  xlab("chlorides")

ggplot(aes(x = density, y = volatile.acidity), data = df) +
  geom_point(aes(color = alcohol.density),
             position = position_jitter(h = 0), alpha = 1/5) +
  scale_colour_brewer(palette = "RdPu")  + theme_dark()+
  scale_x_log10() +
  xlab("density")

ggplot(aes(x = citric.acid, y = density), data = df) +
  geom_point(aes(color =alcohol.density),
             position = position_jitter(h = 0), alpha =1/5) +
  geom_smooth(aes(color = alcohol.density), se = F) +
  scale_colour_brewer(palette = "RdPu")  + theme_dark()


```

From the previous plots i want to explore how the chemical variables interacts with Alcohol.
High alcohol density has low (chlorides, volatile acidity and density)
and free sulfur dioxide increasing with high alcohol density.



```{r echo=FALSE}

ggplot(aes(x = alcohol, y = free.sulfur.dioxide), data = df) +
  geom_point(aes(color = alcohol.quality)) +
  scale_colour_brewer(palette = "RdPu")  + theme_dark()

ggplot(aes(x = alcohol, y = volatile.acidity), data = df) +
  geom_point(aes(color = alcohol.quality), position = position_jitter(h = 0)) +
  scale_colour_brewer(palette = "RdPu")  + theme_dark()

ggplot(aes(x = alcohol, y = density), data = df) +
  geom_point(aes(color = alcohol.quality), position = position_jitter(h = 0)) +
  scale_colour_brewer(palette = "RdPu")  + theme_dark()

```

From the previous plots i want to explore correlation between alcohol quality and alcohol
i found that the wine with high quality has medium alcohol, lower density , high citirc acid and 
low volatile acidity



```{r echo=FALSE}
ggplot(aes(y = pH, x = density), data = df) +
  geom_point(aes(color = df$fixed.acidity), alpha = 0.5) 
```

From the previous chart i found that fixed acidity and density increasing but PH is decreasing. 



# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \investigation. Were there features that strengthened each other in terms of \looking at your feature(s) of interest?

i observed that:
-High alcohol density has low (chlorides, volatile acidity and density)
-free sulfur dioxide increasing with high alcohol density.
- wine with high quality has medium alcohol, lower density , high citirc acid and 
low volatile acidity
-

### Were there any interesting or surprising interactions between features?

High quality wines have lower volatile acidity.

### OPTIONAL: Did you create any models with your dataset? Discuss the \strengths and limitations of your model.

------

# Final Plots and Summary



### Plot One
```{r echo=FALSE, Plot_One}

p1<- qplot(factor(quality), data=df , xlab="Quality", fill=alcohol.quality) + theme_bw()

p1 + ggtitle("number of wines and Quality ") +
  xlab(("Quality")) + ylab("wines#")

```

### Description One
The quality of red wine showes that 80% have a good quality
and the chart shows that is normally distributed around 5 to 6 

### Plot Two
```{r echo=FALSE, Plot_Two}

count_c("free.sulfur.dioxide", 0.5) +
  scale_x_log10()+
  ggtitle("number of wines with free sulfur dioxide")+
  xlab(("free sulfur dioxide %")) + ylab("wine count ")


```

### Description Two
The charts above is normally distributed
the average of free.sulfur.dioxide is 10 



### Plot Three
```{r echo=FALSE, Plot_Three}

ggplot(df, aes(fixed.acidity, pH)) + 
  geom_jitter(alpha=1/5) + 
  stat_smooth(method = "glm", formula = y ~ x)+
  ggtitle("Relationship with fixed.acidity and PH ")+
  xlab(("fixed.acidity %")) + ylab("PH %")


```

### Description Three

The chart above shows that there is negative relationship between (fixed acidity and PH)
 


------

# Reflection

in the begin of the project i try to know the dataset better and i found there is 1,599 observations with 13 variables and i remove x coulomn so there is 12 variables.
and i notice throw the analyze process that the Quality  has strong effect on the variables so i define a new variable "alcohol quality" to see it clearly in chart. 

in the future when i have best quality for wine the graph for alcohol quality
that i ploted will change and the (v.good) will change to become
the highest value insted of medium.
and as a next step i will develop a statistical model for Red wine dataset.



